NN Portfolio Optimierung

### 2024-02-16 paper/Paper.tex
add images and tables for results and conclusions and further research

        \centering
        \large Course 
        \vspace{1mm}       
        \\ \normalsize \textit{Statistical Learning in Finance and Insurance}

@@

    % \newpage \section{Objectives}

@@

    \newpage \section{Data}

@@

@@

                    \includegraphics[scale=0.6]{natural_overview.png}
                    \caption{CIFAR-10 Natural Images (CIFAR) \cite{krizhevsky2009}}
                    \label{cifar-overview}

@@

@@

                    \includegraphics[scale=0.6]{scraped_overview_original.png}
                    \caption{Scraped Graphs (SCP) \cite{sunedition2021}}
                    \label{scraped-overview}

@@

@@

                    \includegraphics[scale=0.6]{norm_460x345_overview.png}
                    \caption{Generated Graphs (GEN)}
                    \label{generated-overview}

@@

                Where W := input volume size, F := receptive field size (kernel), P := amount of zero padding, and S := stride of kernel

@@

                for i = 1, ... , F feature map patches

@@

            In order to understand the results better and have a clear record of all the evaluations, 
            I created a CSV with relevant information. They include the ending prediction weight for each class, 
            the true or false prediction, the file name, the max weight, and the label based on the max.
            With the results readily available, it was easy to see some patterns. 
            In my earlier caveats, it was stated that the normal and log-normal distribution can look quite similar 
            if given the right parameters and that plays out when one looks at 
            how the model misclassified some normal distribution graphs as log-normal. 
            Also, from the earlier section about the generated graphs, 
            I mentioned the so-called ‘duds’ and that comes up in the evaluations as well. 
            All of the uniform distribution graphs misclassified can be considered ‘duds’ in that they appear blank.

            To wrap up my evaluations and to truly assess my generated graphs and models, 
            I needed to use graphs that have not been used in the model previously. 
            I simply gathered 20 graphs of every distribution type from random sources online 
            and then drew one graph of each distribution by hand. I felt this would add some realistic grounding to the final evaluation. 
            Each distribution had 21 graphs, so a total of 84 graphs. 
            The final evaluation accuracy was 59.10 percent and fell in line with my more hopeful expectations. 
            This evaluation is similar to when the scraped graphs where put into CIFAR\_GEN\_1 and if you can remember, 
            the accuracy was only 37.22 percent. So overall, it is a positive development. 
            Just like before the question becomes what was getting misclassified 
            and here is where the results diverged from my original predictions. 
            Almost all of the misclassified graphs were misclassified as log-normal. 
            And if one analyzed this more granularly then one can see 
            that the uniform distribution graphs were overwhelmingly misclassified as log-normal. 
            13 graphs out of the 21 uniform graphs were misclassified as log-normal. 
            It is surprising how 61.90 percent of uniform graphs in the test dataset were misclassified as log-normal 
            since the generated uniform graphs were predicted so accurately in DIST\_153x115\_1.

        \newpage \subsection{Conclusions and Further Research}

            As a whole, I can say that the overall goals were accomplished. 
            I have a model that can successfully identify graphs from natural images 
            with greater than 98 percent accuracy and another model that can successfully identify graphs 
            with four different distribution types with around 93 percent accuracy. 
            Addendum to these results is the caveat from earlier that the generated graphs could be considered as a disjunct subset, 
            if one considers the scraped graphs to be a genuine sample of possible and currently available graphs. 
            The GEN\_SCP\_1’s high accuracy points towards this theory but a large limiting factor is the small sample size -- 
            only 8,000 generated graphs and 7,753 scraped graphs. 
            Some results that surprised me were that scraped graphs in CIFAR\_GEN\_1 were misclassified as natural images 
            around 63 percent of the time. This could indicate 
            that the generated graphs don’t consider the more chaotic nature of graph design. 
            It would be safe to say that the generated graphs are very conservative in appearance. 
            However, this was an original design choice to reduce the amount of variables in the graphs design, 
            like not having two graphs in one image or keeping to the standard of having a complete graph in each image. 
            Many graphs, like the ones in the scraped dataset 
            and the ones I used in the final test evaluation don’t have any axes, labels, or titles. 
            This enforces the idea that many of the images in SCP and many graphs in general are more ‘graph-like’ 
            and not graphs in a very defined sense.
            Another surprise was the difference in how the uniform distribution graphs were getting misclassified.
            Uniform distribution graphs in DIST\_153x115\_1 were classified accurately around 99 percent of the time, 
            but 20 percent of uniform graphs in CIFAR\_SCP\_1 were misclassified as natural images. 
            Why were the graphs easily classified in DIST\_153x115\_1 but not in CIFAR\_SCP\_1? 
            Perhaps graphs with the uniform distribution are unique from the other distribution graphs but not very unique from natural images.

            To conclude I’ll propose some possible changes and improvements to further this line of research. 
            The quickest improvements could simply come from changing the model structure and tweaking the parameters. 
            While I only used one convolutional layer at a time, 
            another common CNN architecture is to stack two convolutional layers before each pooling later. 
            This is strongly encouraged as this allows for more complex features of the input vector to be selected \cite{oshea2015}. 
            Regarding the parameters, I built a second model by doubling the number of filters in the convolution layers. 
            There were minor improvements to the validation accuracy of the DIST\_32x32\_1, 
            but the issue of overfitting is seen after 10 epochs. Secondly, one could use more images to train each model. 
            It would be quite easy to generate 20,000 graphs, but that depends on the quality and ‘generalness’ of the graphs. 
            As seen, the generated graphs used in this experiment were highly distinguishable from the scraped graphs 
            and that leads to the next change, and that is improving the generated graphs. 
            The scripts that created the graphs can be improved to increase variability, 
            but another way is adding more preprocessing layers. TensorFlow has a host of different methods 
            that change brightness, contrast, crop, flip, and rotation, 
            which would do a good job at simulating the chaotic nature of graphs created by people from all over the world. 
            Finally, the experiment could be further by simply expanding the probability distributions used.

            As stated at the beginning, the focus was on identifying and classifying graphs, 
            but it was just the first step in a large picture of identifying, locating, and organizing. 
            To extend a practical use case for this project, one could start using entire reports as input to train the model. 
            In this scenario, pages with graphs on them would be the y target. 
            With the pages identified, they could be returned to the user for quick viewing. 
            From there the model could be expanded to only retrieve the graphs themselves and organize them into types. 
            This would require more advanced techniques, 
            especially those in Explainable AI \cite{kamakshi2023}, 
            to target specific parts on a page and be able to extract more information.

@@

    \begin{table}[htp]
        \begin{center}

            \begin{tabular}{ | l | l | l | l | l | }

                \hline
                \textbf{Ticker}      & \textbf{Name}                             & \textbf{Geschäft} \\
                \hline
                COHR                 & Coherent, Inc.                            & Laser \\
                CTB                  & Cooper Tire \& Rubber Company             & Reifen \\
                EQT                  & EQT Corporation                           & Energie \\
                GOLD                 & Barrick Gold Corporation                  & Bergarbeit \\
                NSEC                 & National Security Group, Inc.             & Versicherung \\
                OTTR                 & Otter Tail Corporation                    & Energie \\
                PEP                  & PepsiCo, Inc.                             & Getränke \\
                SKYW                 & SkyWest, Inc.                             & Fluglinie \\
                SNFCA                & Security National Financial Corporation   & Lebensversicherung \\
                WY                   & Weyerhaeuser Company                      & Abholzung \\
                \hline

        \end{center}
    \end{table}

    \begin{Large} \[ V_p = \sum_{j=1}^{A}\sum_{i=1}^{M}P^T_{i+1} \cdot w_i \] \end{Large}

    \begin{multicols}{2}
        \begin{footnotesize}

            \noindent M = 360 Monate \\
            A = 10 Aktien \\
            P = tägliche Preise von Aktie A im Monat M \\
            w = monatliche Gewicht von Aktie A im Monat M \\
            $P_{1+1}$ = Preise im Februar 1990 von Aktie A \\
            $w_1$ = Gewicht im Januar 1990 von Aktie A

        \end{footnotesize}
    \end{multicols}

>>>>

        \centering
        \large Course 
        \vspace{1mm}       
        \\ \normalsize \textit{Deep Learning Seminar}

@@

@@

    \section{Data}

@@

            \begin{table}[h]

                    \includegraphics[scale=0.6]{natural_overview.png}
                    \caption{CIFAR-10 Natural Images (CIFAR) Deer Examples \cite{krizhevsky2009}}
                    \label{cifar-overview}

@@

@@

            \begin{table}[h]

@@

                    \includegraphics[scale=0.6]{scraped_overview_original.png}
                    \caption{Scraped Graphs (SCP) Bar Chart Examples \cite{sunedition2021}}
                    \label{scraped-overview}

@@

            \begin{table}[h]

                    \includegraphics[scale=0.6]{norm_460x345_overview.png}
                    \caption{Generated Graphs (GEN) Normal Distribution Examples}
                    \label{generated-overview}

@@

@@

                \noindent Where W := input volume size, F := receptive field size (kernel), P := amount of zero padding, 
                and S := stride of kernel

@@

                \noindent for i = 1, ... , F feature map patches

@@

            In order to understand the results better and have a clear record of all the evaluations, 
            I created CSVs with relevant information. They include the ending prediction weight for each class, 
            the true or false prediction, the file name, the max weight, and the label based on the max.
            With the results readily available, it was easy to see some patterns. 
            In my earlier caveats, it was stated that the normal and log-normal distribution can look quite similar 
            if given the right parameters and that plays out when one looks at 
            how the model misclassified some normal distribution graphs as log-normal. 
            Also, from the earlier section about the generated graphs, 
            I mentioned the so-called ‘duds’ and that comes up in the evaluations as well. 
            All of the uniform distribution graphs misclassified can be considered ‘duds’ in that they appear blank.

                \scalebox{0.75}{

                    \begin{tabular}{l | l | l | l | l | l | l | l | l}
                        \textbf{exp}         & \textbf{lognorm}     & \textbf{norm}         & \textbf{unif}       & \textbf{max}       & \textbf{file\_name}         & \textbf{actual}  & \textbf{label}   & \textbf{prediction} \\
                        -4.1661463  & 8.161812    & 0.9788155    & -10.229221 & 8.161812  & exp\_12.jpeg       & exp     & lognorm & False      \\
                        3.8511171   & 4.496506    & -3.405139    & -5.2995405 & 4.496506  & exp\_2.jpeg        & exp     & lognorm & False      \\
                        5.872565    & 11.971514   & -6.275291    & -12.128517 & 11.971514 & exp\_20.jpeg       & exp     & lognorm & False      \\
                        9.207311    & 10.764248   & -11.954131   & -5.3854074 & 10.764248 & exp\_3.jpeg        & exp     & lognorm & False      \\
                        -5.917483   & 15.367903   & -4.7242074   & -5.200668  & 15.367903 & exp\_5.jpeg        & exp     & lognorm & False      \\
                        6.1337843   & 6.400875    & -5.021023    & -7.4242744 & 6.400875  & exp\_8.jpeg        & exp     & lognorm & False      \\
                        0.51164746  & 8.585008    & -3.2878444   & -6.293257  & 8.585008  & exp\_9.jpeg        & exp     & lognorm & False      \\
                        -2.7664876  & 8.546496    & 0.33398134   & -9.835137  & 8.546496  & exp\_hand\_0.jpeg  & exp     & lognorm & False      \\
                        -13.257631  & 5.875923    & -2.4565024   & 7.38458    & 7.38458   & lognorm\_0.jpeg    & lognorm & unif    & False      \\
                        3.7302907   & 2.0135944   & -0.048096493 & -11.51519  & 3.7302907 & lognorm\_5.jpeg    & lognorm & exp     & False      \\
                        -14.977435  & 3.678837    & 1.5030111    & 1.7347182  & 3.678837  & norm\_1.jpeg       & norm    & lognorm & False      \\
                        -0.31633773 & 5.5405626   & 1.6448251    & -18.00231  & 5.5405626 & norm\_11.jpeg      & norm    & lognorm & False      \\
                        -14.40909   & 10.438942   & 7.7604384    & -26.327517 & 10.438942 & norm\_13.jpeg      & norm    & lognorm & False      \\
                        -5.3496103  & 7.2749767   & 1.1707792    & -6.384265  & 7.2749767 & norm\_14.jpeg      & norm    & lognorm & False      \\
                        -6.992462   & 7.189061    & 4.9432883    & -15.288852 & 7.189061  & norm\_15.jpeg      & norm    & lognorm & False      \\
                        -4.783489   & 4.332031    & 3.8387465    & -14.778937 & 4.332031  & norm\_19.jpeg      & norm    & lognorm & False      \\
                        -23.318886  & 12.883771   & -0.8641091   & 5.143599   & 12.883771 & norm\_4.jpeg       & norm    & lognorm & False      \\
                        -25.816198  & 16.998684   & 4.5914965    & -9.728744  & 16.998684 & norm\_8.jpeg       & norm    & lognorm & False      \\
                        -6.6115365  & 7.848427    & 0.57640886   & -4.26034   & 7.848427  & norm\_9.jpeg       & norm    & lognorm & False      \\
                        -0.7095418  & 1.1841375   & -0.53598833  & -1.6236649 & 1.1841375 & unif\_0.jpeg       & unif    & lognorm & False      \\
                        -4.572587   & 5.3358355   & 0.4110088    & -3.925213  & 5.3358355 & unif\_1.jpeg       & unif    & lognorm & False      \\
                        -2.4820209  & -0.05105117 & 2.7819424    & -4.3270063 & 2.7819424 & unif\_10.jpeg      & unif    & norm    & False      \\
                        -4.820385   & 9.742123    & -7.255406    & 6.477165   & 9.742123  & unif\_11.jpeg      & unif    & lognorm & False      \\
                        -0.5606266  & 6.14713     & -2.2170813   & -3.020845  & 6.14713   & unif\_12.jpeg      & unif    & lognorm & False      \\
                        1.1152176   & 1.1827534   & -0.67093164  & -5.319949  & 1.1827534 & unif\_13.jpeg      & unif    & lognorm & False      \\
                        -0.19624099 & 7.653762    & -2.4959579   & -3.569571  & 7.653762  & unif\_14.jpeg      & unif    & lognorm & False      \\
                        -0.46984562 & 5.252478    & -4.9778605   & 1.5037757  & 5.252478  & unif\_16.jpeg      & unif    & lognorm & False      \\
                        -0.20320779 & 4.1991606   & 1.5740888    & -10.120164 & 4.1991606 & unif\_17.jpeg      & unif    & lognorm & False      \\
                        -2.3884668  & 3.9697368   & -0.6534233   & -2.8202808 & 3.9697368 & unif\_2.jpeg       & unif    & lognorm & False      \\
                        1.6306777   & 6.3168035   & -0.7305184   & -8.0671625 & 6.3168035 & unif\_3.jpeg       & unif    & lognorm & False      \\
                        -5.6656113  & 4.946749    & 1.8187       & -6.4847775 & 4.946749  & unif\_4.jpeg       & unif    & lognorm & False      \\
                        -4.4202156  & 3.0252044   & 1.0488026    & -2.0555344 & 3.0252044 & unif\_5.jpeg       & unif    & lognorm & False      \\
                        5.331371    & 4.629284    & -3.9633603   & -6.628367  & 5.331371  & unif\_6.jpeg       & unif    & exp     & False      \\
                        -4.064364   & 4.5016537   & -3.487114    & 3.3282218  & 4.5016537 & unif\_8.jpeg       & unif    & lognorm & False      \\
                        -6.592189   & -0.6548197  & 2.7110033    & -1.7038764 & 2.7110033 & unif\_9.jpeg       & unif    & norm    & False      \\
                        0.5345101   & 0.08001919  & 0.43601054   & -6.1528854 & 0.5345101 & unif\_hand\_0.jpeg & unif    & exp     & False      \\

                }

                \caption{All Misclassified Images in Final Test Evaluation}
                \label{all-mis-test-eval}

            To wrap up my evaluations and to truly assess my generated graphs and models, 
            I needed to use graphs that have not been used in any model previously. 
            Visual \ref{all-mis-test-eval} contains all images in the final test evaluation that were misclassified.
            To get this, I simply gathered 20 graphs of every distribution type from random sources online 
            and then added one graph of each distribution drawn by hand. 
            I felt this would add some realistic grounding to the final evaluation. 
            Each distribution had 21 graphs, so a total of 84 graphs. 
            The final evaluation accuracy was 59.10 percent and fell in line with my more hopeful expectations. 
            This evaluation is similar to when the scraped graphs where put into CIFAR\_GEN\_1 and if you can remember, 
            the accuracy was only 37.22 percent. So overall, it is a positive development. 
            Just like before the question becomes what was getting misclassified 
            and here is where the results diverged from my original predictions. 
            Almost all of the misclassified graphs were misclassified as log-normal,
            which can be seen in the 'label' column in Visual \ref{all-mis-test-eval}. 
            And if one analyzed this more granularly then one can see 
            that the uniform distribution graphs were overwhelmingly misclassified as log-normal. 
            13 graphs out of the 21 uniform graphs were misclassified as log-normal. 
            It is surprising how 61.90 percent of uniform graphs in the test dataset were misclassified as log-normal 
            since the generated uniform graphs were predicted so accurately in DIST\_153x115\_1.
            I can conclude for now that the model needs more diverse training images to account for the variety in graphs found online
            and more images to train on.

        \subsection{Conclusions and Further Research}

            As a whole, I can say that the main goals were accomplished. 
            I have a model that can successfully identify graphs from natural images 
            with greater than 98 percent accuracy and another model that can successfully identify my generated graphs 
            with four different distribution types with around 93 percent accuracy. 
            However, when more diverse graphs are included this drops down to 59.10 percent as shown in the 
            final test evaluation and in Visual \ref{all-mis-test-eval}.
            Addendum to these results is the caveat from earlier that the generated graphs could be considered as a disjunct subset, 
            if one considers the scraped graphs to be a genuine sample of possible and currently available graphs. 
            The GEN\_SCP\_1’s high accuracy points towards this theory but a large limiting factor is the small sample size -- 
            only 8,000 generated graphs and 7,753 scraped graphs. 
            Some results that surprised me were that scraped graphs in CIFAR\_GEN\_1 were misclassified as natural images 
            around 63 percent of the time. This could indicate 
            that the generated graphs don’t consider the more chaotic nature of graph design -- 
            this is substantiated by the final test evaluation as well.
            It would be safe to say that the generated graphs are very conservative in appearance. 
            However, this was an original design choice to reduce the amount of variables in the graphs design, 
            like not having two graphs in one image or keeping to the standard of having a complete graph in each image. 
            Many graphs, like the ones in the scraped dataset 
            and the ones I used in the final test evaluation, don’t have any axes, labels, or titles. 
            This enforces the idea that many of the images in SCP and many graphs in general are more ‘graph-like’ 
            and not graphs in a very defined sense.
            Another surprise was the difference in how the uniform distribution graphs were getting misclassified.
            Uniform distribution graphs in DIST\_153x115\_1 were classified accurately around 99 percent of the time, 
            but 20 percent of generated uniform graphs in CIFAR\_SCP\_1 were misclassified as natural images. 
            Why were the graphs easily classified in DIST\_153x115\_1 but not in CIFAR\_SCP\_1? 
            Perhaps graphs with the uniform distribution are unique from the other distribution graphs 
            but not very unique from natural images. The final test evaluation bucked this trend by
            classifying uniform distribution as log-normal for a majority of the graphs.
            However, in the final test evaluation almost all of the misclassified images were
            classified as log-normal, so this might point to an issue with the training images.

            \begin{table}[h]

                    \includegraphics[scale=0.6]{DIST_32x32_2_HIST_RESULTS.png}
                    \caption{DIST\_32x32\_2 Training and Validation Accuracy and Loss}
                    \label{dist-32-32-2-val-loss}

            To conclude I’ll propose some possible changes and improvements to further this line of research. 
            The quickest improvements could simply come from changing the model structure and tweaking the parameters. 
            While I only used one convolutional layer at a time, 
            another common CNN architecture is to stack two convolutional layers before each pooling later. 
            This is strongly encouraged as this allows for more complex features of the input vector to be selected \cite{oshea2015}. 
            Regarding the parameters, I built a second model by doubling the number of filters in the convolution layers. 
            The training and validation accuracy and loss of DIST\_32x32\_2
            are shown in Visual \ref{dist-32-32-2-val-loss} --
            the 2 indicates the second model used.
            There were minor improvements to the validation accuracy of the DIST\_32x32\_1, 
            but the issue of overfitting is seen after 10 epochs. Secondly, one could use more images to train each model. 
            It would be quite easy to generate 20,000 graphs, but that depends on the quality and ‘generalness’ of the graphs. 
            The generated graphs used in this experiment were highly distinguishable from the scraped graphs 
            and that leads to the next change, and that is improving the generated graphs. 
            The scripts that created the graphs can be improved to increase variability, 
            but another way is adding more preprocessing layers. TensorFlow has a host of different methods 
            that change brightness, contrast, crop, flip, and rotation, 
            which would do a good job at simulating the chaotic nature of graphs created by people from all over the world. 
            Finally, the experiment could be further by simply expanding the probability distributions used.

            As stated at the beginning, the focus was on identifying and classifying graphs, 
            but it was just the first step in a large picture of identifying, locating, and organizing. 
            To extend a practical use case for this project, one could start using entire reports as input to train the model. 
            In this scenario, pages with graphs on them would be the y target. 
            With the pages identified, they could be returned to the user for quick viewing. 
            From there the model could be expanded to only retrieve the graphs themselves and organize them into types. 
            This would require more advanced techniques, 
            especially those in the field of explainable AI \cite{kamakshi2023}, 
            to target specific parts on a page and be able to extract more information.

@@
